---
title: "PA_1template.rmd"
author: "Parichay"
date: "20/06/2020"
output: html_document
---

Creating variables Activity and activity 1, the latter having omitted NA values
```{r,  echo=TRUE}
activity <- read.csv("./activity.csv")
activity1 <- na.omit(activity)
```

Creating histogram of aggregate number of steps per day and calculating mean and median of the data.
```{r, echo=TRUE}
steps_per_day<- aggregate(steps~date, activity1, sum)
hist(steps_per_day$steps, main = "Histogram of total number of steps per day", xlab = "Steps per day")
mean(steps_per_day$steps)
median(steps_per_day$steps)
```
Average daily pattern
```{r, echo=TRUE}
steps_per_interval<- aggregate(steps~interval, activity1, sum)
plot(steps_per_interval$steps,type="l", main = "Histogram of total number of steps per interval", xlab = "Interval", ylab = "steps")
steps_per_interval$interval[which.max(steps_per_interval$steps)]
```
Histogram of the total number of steps taken each day after imputing missing values and the mean and median total number of steps taken per day.
```{r, echo=TRUE}
avg_steps<- aggregate(steps~interval,activity, mean)
imp <- activity # new dataset called imp
 for (i in avg_steps$interval) {
    imp[imp$interval == i & is.na(imp$steps), ]$steps <- 
    avg_steps$steps[avg_steps$interval == i]
 }
total_step_imp <- aggregate(steps ~ date, imp, sum)
hist(total_step_imp$steps, breaks =20, 
     main = "Total Number of Steps Taken Each Day (Imputed)",
     col = "grey", border = "white", xlab = "Step")
mean(total_step_imp$steps)
median(total_step_imp$steps)
```
Difference pattern in weekdays and weekends

```{r, echo=TRUE}
# function to differentiate weekday and weekend
weekday <- function(date_val) {
   wd <- weekdays(as.Date(date_val, '%Y-%m-%d'))
    if  (!(wd == 'Saturday' || wd == 'Sunday')) {
        x <- 'Weekday'
    } else {
     x <- 'Weekend'
 }
     x
}
activity$day_type <- as.factor(sapply(activity$date, weekday))
library(ggplot2)
steps_per_day_impute <- aggregate(steps ~ interval+day_type, activity, mean)
# Create the plot
plt <- ggplot(steps_per_day_impute, aes(interval, steps)) +
     geom_line(stat = "identity", aes(colour = day_type)) +
     theme_gray() +
     facet_grid(day_type ~ ., scales="fixed", space="fixed") +
     labs(x="Interval", y=expression("No of Steps")) +
     ggtitle("No of steps Per Interval by day type")
print(plt)
```

